---
title: "EFASH"
author: "Youngseok Kim"
output: workflowr::wflow_html
---

## Examples

#### Gaussian-Gaussian case

In case of Gaussian, we have $G(x) = x^2/2$, $\Delta_G(x\|y) = (x-y)^2/2$, $D_G(x\|y) = (x-y)^2/2 - x^2/2$ and 

$$
H(\nu,t) = e^{\nu^2/2} \sqrt{t/2\pi}
$$

since

$$
\mu_j|\nu_k,t_k \sim N(\mu_k,1/t_k)
$$

#### Poisson-Gamma case

In case of Poisson-Gamma, we have $G(x) = x(\log x-1)$, $\Delta_G(x\|y) = x(\log x- \log y) - x + y$, $D_G(x\|y) = -x \log y + y$ and

$$
H(\nu,t) = t^{t\nu + 1}\Gamma(t\nu+1)^{-1}
$$

$$
\mu_j|\nu_k,t_k \sim \textrm{Gamma}(t_k\nu_k+1,1/t_k)
$$

and

$$
\exp\left[t\nu \log\mu - t\mu\right] = \mu^{t\nu}e^{-t\mu}
$$

## Binomial-Beta case

Let $n$ be the number of Bernoulli trial. In case of Beta, we have $G(x) = x \log \frac{x}{n} + (n-x) \log \frac{n-x}{n}$, $\Delta_G(x\|y) = x\log \frac{x}{y} + (n-x)\log \frac{n-x}{n-y}$, $D_G(x\|y) = - x\log \frac{y}{n} - (n-x) \log \frac{n-y}{n}$ and

$$
H(\nu,t) = \frac{\Gamma(nt+2)}{\Gamma(nt\nu + 1)\Gamma(nt(1-\nu)+1)}
$$

since

$$
\mu_j|\nu_k,t_k \sim \textrm{Beta}(nt_k\nu_k+1,nt_k(1-\nu_k)+1)
$$

#### Dirichlet case


```{r}
H_partition_function <- function(family = "normal", G = NA, D_G = NA){
  
  if(family == "normal" || family == "gaussian"){
    H <- function(nu,t){
      return( sqrt(t/2/pi) )
    }
    h <- function(nu,t){
      return( log(t)/2 - log(2*pi)/2 )
    }
  }
  
  if(family == "gamma"){
     H <- function(nu,t){
      return( t^(t * nu + 1) * gamma(t * nu + 1)^(-1) )
     }
     h <- function(nu,t){
      return( (t * nu + 1) * log(t) - lgamma(t * nu + 1))
    }
  }
  
  if(family == "beta"){
    
  }
  
  if(family == "dirichlet"){
    
  }
}
```


```{r}
H_partition_manual(G = NA, D_G = NA){
  if(G == NA){
    
  }
}
```
